The head moves by rotation and translation, or
yaw, pitch, roll, left/right, up/down or forward/back, which sums up to six degrees of freedom(6DOF).
Head tracking refers to the estimation of rotation, translation or both, usually based on 
optical sensors like cameras and infrared lights.

TrackIR\ref{trackir} is a commercial product by NaturalPoint for headtracking, costing 700-1050NOK, which provides 6DOF head tracking.
The algorithm for head tracking is stated by TrackIR to be proprietary and not disclosed to the public. TrackIR only supports the Microsoft Windows operating system.

FreeTrack is a free optical motion tracking application for Microsoft Windows\ref{freetrack}. The software uses
a webcam for head tracking with the user wearing for example a cap inrared lights. FreeTrack provides 2DOF or 6DOF head tracking.
The algorithms for head tracking used in FreeTrack are Dementhon's four point algorithm\ref{dementhon}, and Alter's three point algorithm\ref{alter}

The typical use of the two software packages is playing flight- or racing simulator games
to look around in the cockpit or out the side windows by rotating the head. 
This use of head tracking does not use accurate 3D-perspective projection, partly because it is
necessary for the virtual eye to move or rotate much more than the players head in order for the player to always see the screen, referred to as movement scaling.

The solution proposed here calculates head translation and thus provides 3DOF head tracking. 
This is sufficient since virtual reality, by accurate 3D perspective projection, is only dependant on head translation.
However, depending on which mode the system is in, which can be either one optical sensor mode or two optical sensors mode,
the user will still have 5DOF and 6DOF movement respectively. 

Two demonstration applications showing the use of head tracking and accurate 3D perspective projection are included in the solution.
The first demonstration application shows how an illusion of objects breaking the surface in the screen can be achieved using head tracking,
giving a very immersive experience. The second demonstration is an equally or even more immersive game, where the player must avoid
obstacles hurdled towards him, by ducking, jumping and dodging side-to-side.
